\begin{thebibliography}{}
\bibitem{plasma} Freidberg, J.P.: \textit{Plasma Physics and Fusion Energy}. New York: Cambridge University Press, 2007.
\bibitem{CHEN} Chen, F.F.: \textit{Introduction to plasma physics}. New York: Plenum Press, 1974.
\bibitem{GUO} Guo, Z.: \textit{Notes on the Magnetic Mirror}. Los Alamos, 2009.
\bibitem{idealMHD} Freidberg, J.P.: \textit{Ideal Maghetohydrodynamics}. New York: Plenum Press, 1987.
\bibitem{magnetic_mirror} Post, R.F.: \textit{The magnetic mirror approach to fusion}. Nuclear Fusion, 27 (10), 1579-1739, 1987.
\bibitem{fluxcoord} D'haeseller, W.D., Hitchon, W.N.G., Callen, J.D. and Shohet, J.L.: \textit{Flux Coordinates and Magnetic Field Structure: A guide to a Fundamental Tool of Plasma Theory}. Berlin: Springer-Verlag, 1991.

\bibitem{QUARTERONI} Quarteroni, A.: \textit{Numerical Models for Differential Problems}. Milan: Springer-Verlag, 2009.
\bibitem{CHQZ06} Canuto, C., Houssain, M.Y., Quarteroni, A., and Zang, T.A.: \textit{Spectral Methods: Fundamentals in Single Domains}. Berlin: Springer-Verlag, 2006.
\bibitem{CHQZ07} Canuto, C., Houssain, M.Y., Quarteroni, A., and Zang, T.A.: \textit{Spectral Methods: Evolution to Complex Geometries and Applications to Fluid Dynamics}. Berlin:
Springer-Verlag, 2007.
\bibitem{KARNIADAKIS} Karniadakis, G.E., and Sherwin, S.: \textit{Spectral/hp Element Methods for Computational Fluid Dynamics}. 2nd ed. New York: Oxford University Press, 2005.
\bibitem{HENDERSON} Henderson, R.D.: \textit{Adaptive spectral element method for turbulence and transition}. In T.J. Barth \& H. Deconinck (Eds.), High-Order Methods for Computational Physics (pp. 225-326). Berlin: Springer-Verlag, 1999.
\bibitem{SERRA} Badiale, M., Serra, E.:\textit{Semilinear Elliptic Equations for Beginners}. London: Springer-Verlag, 2011.

\bibitem{gmsh_man}Geuzaine, C., Remacle, J.F.: \textit{GMSH reference manual. Version 2.6.} http://geuz.org/gmsh/, 2012.
\bibitem{gmsh_art}Geuzaine, C., Remacle, J.F.: \textit{GMSH: a three-dimensional finite element mesh generator with built-in pre-and-post processing facilities.} International Journal for Numerical Methods in Engineering, Vol. 79, issue 11, pages 1309-1331, 2009.
\bibitem{wolfram_struct}Uznanski, D.: \textit{Grid.} MathWorld--A Wolfram Web Resource, created by Weisstein E.W., http://mathworld.wolfram.com/Grid.html
\bibitem{metis}Karypis, G.: \textit{METIS. A Software Package for Partitioning Unstructured Graphs, Partitioning Meshes, and Computing Fill-Reducing Orderings of Sparse Matrices Version 5.0.} http://glaros.dtc.umn.edu/gkhome/, 2011.
\bibitem{parmetis}Karypis, G., Schloegel, K.: \textit{PARMETIS. Parallel Graph Partitioning and Sparse Matrix Ordering Library Version 4.0} http://glaros.dtc.umn.edu/gkhome/, 2011.
\bibitem{mpi}Various Authors: \textit{MPI: A Message-Passing Interface Standard. Version 3.0.} http://www.mcs.anl.gov/research/projects/mpi/, 2012.
\bibitem{gsl}Galassi, M., Davies, J., Theile, J., Gough, B., Jungman, G., Alken, B., Booth, M., Rossi, F.: \textit{GNU Scientific Library Reference Manual. Edition 1.15} http://www.gnu.org/software/gsl/, 2011.
\bibitem{petsc}Balay, S., Brown, J., Buschelman, K.,  Eijkhout, V., Gropp, W., Kaushik, D., Knepley, M., Curfman McInnes, L., Smith, B., Zhang, H.: \textit{PETSc Users Manual. Revision 3.3} http://www.mcs.anl.gov/petsc/, 2012.

\bibitem{cpp_primer}Lippman, S., Lajoie, J., Moo, B.: \textit{C++ Primer}. Addison-Wesley Professional; 5 edition, 2012.

\end{thebibliography}
